---
title: "Linear Regressions"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

# What is a simple linear regression? 


```{r}
library(tidyverse)
theme_set(theme_light())
```

```{r}
adversiting <- read_csv("ISLR/ISLR-Auto/Advertising.csv")

adversiting <- adversiting %>% 
  select(-X1)

adversiting_tided <- adversiting %>% 
  select(-X1) %>% 
  gather(channel, budget , -Sales) %>% 
  rename_with(tolower)

```

```{r}
library(modelr)

TV_data <- adversiting_tided %>% 
  filter(channel == "TV")

smod <- lm(sales ~ budget, data = TV_data)
  
grid <- TV_data %>% 
  add_predictions(smod)


TV_data %>% ggplot(aes(budget, sales)) +
  geom_point(colour = "red") +
  geom_smooth(method = "lm", se = FALSE) + 
  geom_segment(data = grid, 
               aes(xend = budget, yend = pred),
               colour = "grey",
               alpha = 0.5)
```


## Estimate the coefficients 


```{r}
library(broom)
tidy(smod)
```

p-value is relatively big so we safely conclude that coefficients are different from 0. 

## Accuracy of model 

* Residual standard error (RSE): lack of fit
* Rsquared: proportion of variability in Y that can be explained using X
```{r}
glance(smod)
```

In general, actual sales in each market deviate from the true regression line by approximately 3,260 units, on average. Two-thirds of the variability in sales is explained by a linear regression on TV.


# What is a multiple linear regression? 

```{r}
mmod <- lm(Sales ~ TV + Radio + Newspaper, data = adversiting)
```

* Coefficients 


``` {r}
tidy(mmod)
```

* Statistics
``` {r}
glance(mmod)
```

### Questions

* Is There a Relationship Between the Response and Predictors?
* Model fit
* Prediction and confident interval 

## Qualitative Preditors

```{r}
credit <- read_csv("ISLR/ISLR-Auto/Credit.csv")

credit <- credit %>% 
  select(-X1) %>% 
  rename_with(tolower)
```


```{r}
lm(balance ~ gender, data = credit)

lm(balance ~ ethnicity, data = credit)
```

## Extension of Linear model 

Remove addidive assumptions by add interaction term 

```{r}
categorical_mod <- credit %>% 
  lm(balance ~ income + student, data = .)

p1 <- credit %>% 
  add_predictions(categorical_mod) %>% 
  ggplot(aes(income, pred, colour = student)) + 
  geom_line(size = 1) + 
  labs(
    x = "Income",
    y = "Balance Account",
    title = "Linear regression of balance on income for studenst and non-students",
    subtitle = "Without interaction term"
  )
p1
```


```{r}
p2 <- credit %>% 
  ggplot(aes(income, balance, colour = student)) + 
  geom_smooth(method = "lm", se = FALSE) + 
  labs(
    x = "Income",
    y = "Balance Account",
    title = "Linear regression of balance on income for studenst and non-students",
    subtitle = "With interaction term"
  )
p2
```

```{r}
library(gridExtra)
grid.arrange(p1, p2, ncol = 2)
```

## Non-linear relationship

Approach: *polynominal regression*

```{r}
auto <- as_tibble(ISLR::Auto)
mod1 <- auto %>% 
  lm(mpg ~ poly(horsepower,1), data = .) 

mod2 <- auto %>% 
  lm(mpg ~ poly(horsepower,2), data = .) 

mod5 <- auto %>% 
  lm(mpg ~ poly(horsepower,5), data = .)

predictions <- auto %>% 
  gather_residuals(Linear = mod1, Degree2 = mod2, Degree = mod5) %>% 
  gather_predictions(Linear = mod1, Degree2 = mod2, Degree5 = mod5)


auto %>% 
  ggplot(aes(horsepower, mpg)) + 
  geom_point(size = 2, shape = 1, alpha = 0.5) + 
  geom_line(data = predictions,
            aes(horsepower, pred, colour = model))
```

## Potential Problems 

### Non-linearity

```{r}
predictions %>% 
  ggplot(aes(pred, resid)) +
  geom_point(size = 2, shape = 1, alpha = 0.2) +
  geom_smooth(se = FALSE) +
  facet_wrap(~ model, scales = "free")
```

