---
title: "Models with broom and purrr"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
theme_set(theme_light())
```


```{r}
library(gapminder)
gapminder %>% 
  count(country)

nz <- gapminder %>%
  filter(country == "New Zealand")

full <- nz %>% 
  ggplot(aes(year, lifeExp)) + 
  geom_line() + 
  ggtitle("Full data = ") 

```

Linear regression 

```{r}
nz_mod <- nz %>% 
  lm(lifeExp ~ year, data = .)

library(modelr)
grid <- nz %>% 
  data_grid(year) %>% 
  add_predictions(nz_mod) 

linear <- grid %>% 
  ggplot(aes(year, pred)) + 
  geom_line() + 
  ggtitle("Linear")
```

```{r}
resid_pattern <- nz %>% 
  add_residuals(nz_mod) %>% 
  ggplot(aes(year, resid)) + 
  geom_ref_line(h = 0, colour = "grey") +
  geom_line() + 
  ggtitle("Resid")
```

```{r}
library(gridExtra)
grid.arrange(full, linear, resid_pattern, ncol = 3)
```


### How can we plot these for all countries? 

Instead of repeating for each variable, we repeat of each observation of a variable

```{r}
by_country <- 
  gapminder %>% 
  group_by(country, continent) %>% 
  nest()
```

**Step1** model across the dataframe

```{r}
country_model <- function(df) {
  lm(lifeExp ~ year, data = df)
}
```

```{r}
by_country <- by_country %>% 
  mutate(model = map(data, country_model))

by_country <- by_country %>% 
  mutate(resids = map2(data, model, add_residuals))

resids <- by_country %>% 
  unnest(resids)
```

Finally we get the data frame.
Now we can plot 
```{r}
resids %>% 
  ggplot(aes(year, resid, group = country)) +
  geom_line(alpha = 0.3) + 
  facet_wrap(~continent)
```

It seems our model doesn't fit the Africa well which suggests something interesting there

### Out of residuals, model quality is another approach 

```{r}
library(broom)
glance(nz_mod)
```


```{r}
model_quality <- by_country %>% 
  mutate(glance = map(model, glance)) %>% 
  unnest(glance) %>% 
  select(-is.list) %>% 
  arrange(r.squared)
```

Most of the worst models are in Africa 

```{r}
model_quality %>% 
  ggplot(aes(continent, r.squared)) + 
  geom_jitter(width = 0.5)
 
```

Zoom in R-square that is small
```{r}
bad_fit <- model_quality %>% 
  filter(r.squared < 0.25)

gapminder %>% 
  semi_join(bad_fit, by = "country") %>% 
  ggplot(aes(year, lifeExp, colour = country)) + 
  geom_line()
```

