---
title: 'Data Transformation with dplyr '
author: thinh
date: '2020-04-04'
slug: data-transformation-with-dplyr
categories:
  - rfds
tags:
  - R
comments: no
cover: ''
imgs: []
justify: no
lastmod: '2020-04-04T14:35:07+07:00'
license: ''
readingTime: yes
single: no
output: 
  html_document: 
    theme: cosmo
    toc: yes
---
# Prerequisites
```{r}
library(tidyverse)
library(nycflights13)
```
## nyclights13
 Data types of the variables:
 
- `int`: integers
- `dbl`: double/real numbers 
- `chr`: character/string
- `dttm`: date + time
- `lgl`: logical. `TRUE` OR `FALSE`
- `fctr`: factors -> categorical data with fixed possible values
- `date`: dates

## dplyr Basics

Key functions: 

- pick observations - `filter()` - deal with rows 
- pick variables - `select()` - deal with columns 
- create new variable - `mutate()` - insert a new columns 
- reorder the rows - `arrange()` 
- collapse values into a single summary - `summarize()`
 
Used in conjunction with `group_by`

How it works: data.frame -> arguments -> result

# Filter rows with filter() 
- subset observations based on their values. 

```{r}
filter(flights, month == 1, day == 1)
```

## Comparisons & Logical Operators
- Comparisons: `>`, `<`, `=`, `>=`,`<=`, `==`, `!=`
- Logical operators: `&` , `|`
- To simplify `x == 1 | x == 2`, we can apply `x %in% c(1,2)` 

```{r}
filter(flights, month %in% c(1,2))
```

## Missing values 
```{r}
# not available
NA
```
determine a observation is NA

```{r}
x <- NA
is.na(x)
```

# Arrange rows with arrange()
- similar to `filter()`
```{r}
arrange(flights, year, month, day)
```

- use desc(x) to reorder in descending order

```{r}
arrange(flights, desc(month))
```

# Select columns with select()
- to zoom quickly to the subset
```{r}
select(flights, year, month, day)
```

- select all columns between a and b 

```{r}
select(flights, year:day)
```

- select all columns except a and b 

```{r}
select(flights, -(year:day))
```

- reorder columns 

```{r}
select(flights, day, everything()) # put day to the first column and keep same for everything
```


# Add new variable with mutate()
- `mutate()` always add a new variable at the end of the dataset

```{r}
flights_test <- select(flights, 
                       year:day, 
                       ends_with("delay"), 
                       distance, 
                       air_time)
mutate(flights_test, 
       gain = arr_delay - dep_delay,# new variable 1 
       speed = distance/ air_time * 60, # new variable 2 
       hours = air_time/60, # new variable 
       gain_per_hour = gain/hours # new variable by using new variable 1&3
       )
```

- show only new varialbe then we use `transmute`

```{r}
transmute(flights_test, 
       gain = arr_delay - dep_delay,# new variable 1 
       speed = distance/ air_time * 60, # new variable 2 
       hours = air_time/60, # new variable 
       gain_per_hour = gain/hours # new variable by using new variable 1&3
       )
```

## Useful creation with mutate()
- *Arithmetic operator*: `+, -, /, *, ^`
- *Modular arithmetic*: %/% (integer), %/% (remainder)
- *Log*: log()
- *Offsets*: lag(), lead()
- *Cumulative and rolling aggregates*: `cumsum()`, `cumprod()`, `cumnin()`, `cummax()`, `cummean()`
- *Logical comparisons*: `<, <=, >, >=, !=, ==`
- *Ranking*: `min_rank()`, `row_number()`, `dense_rank()`, `percent_rank()`

# Grouped Summary with summarize()
- collapse data into a single row

```{r}
# average dep delay
summarise(flights, delay = mean(dep_delay, na.rm = TRUE)) # drop missing value 
```
- combining with `group_by` makes it more useful 
```{r}
# average dep delay per day
by_day <- group_by(flights, year, month, day)
summarise(by_day, delay = mean(dep_delay, na.rm = TRUE)) 
```

## Combining Multiple Operators with the Pipe
- Explore the relationship between distance and average delay for each location. 

```{r}
delay <- flights %>%
  group_by(dest) %>% 
  summarise(
    count = n(),
    dist = mean(distance, na.rm = TRUE),
    delay = mean(arr_delay, na.rm = TRUE)
  ) %>% 
  filter(count > 20, dest != "HML")
```
 
- pronounce `%>%`  as `then`. Behind the scences, `x %>% f(y)` is equivalent to `f(x,y)`. `x %>% f(y) %>% g(z)` is equivalent to `g(f(x,y), z)`

## Missing Values
If we don't put `na.rm` into the `summarise()`, then the result is NA as expected 
```{r}
flights %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay)) 
```

In this case, the missing values represent for the cancelled flights. We can remove them for the dataset as 

```{r}
not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay) )

not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay))
```

# Counts 
-  a count `n()` or a count of non missing values `sum(!is.na(x))`

**Question**: WHich planes have the highest average delays? 
Planes are defined by their tail number

```{r}
delays <- not_cancelled %>% 
  group_by(tailnum) %>% 
  summarise(delay = mean(arr_delay))

ggplot(data = delays, 
       mapping = aes(x = delay)) + 
  geom_freqpoly(binwidth = 10)
```

Number of flights versus average delay 

```{r}
delays <- not_cancelled %>% 
  group_by(tailnum) %>% 
  summarise(delay = mean(arr_delay),
            n = n() # number flight
            )

ggplot(data = delays, 
       mapping = aes(x = n, y = delay)) +
  geom_point(alpha = 1/10)
```

There is much variations when there are few flights. So we filter out the group with the smallest numbers of observation 

```{r}
delays %>%
  filter(n > 25) %>% 
  ggplot(mapping = aes(x = n, y = delay)) +
  geom_point(alpha = 1/10)
```

