---
title: Dates and Times with lubridate
author: thinh
date: '2020-04-08'
slug: dates-and-times-with-lubridate
categories:
  - rfds
tags:
  - R
comments: no
cover: ''
imgs: []
justify: no
lastmod: '2020-04-08T22:23:40+07:00'
license: ''
readingTime: yes
single: no
toc: yes
---
# Prerequisites

```{r}
library(tidyverse)
library(lubridate)
library(nycflights13)
```


# Creating Date/Times
Three type of date/time data

- *date*. <date>
- *time*. <time>
- *date-time*. <dttm>

```{r}
today()
now()
```

Three ways to create a time 

- string 
- date-time components
- existing date/time object

## From strings

```{r}
ymd(20200330) # or ymd("2020-03-30")
```

```{r}
ymd_hms(now())
```

# From Individual Components

```{r}
flights %>% 
  select(year, month, day, hour, minute)
```

To make date time 

```{r}
flights %>% 
  select(year, month, day, hour, minute) %>% 
  mutate(departure = make_datetime(year, month, day, hour, minute))
```

## From Other Type

```{r}
as_datetime(today())
```

```{r}
as_date(now())
```

# Date-Time Components
## Getting components
- `year()`, 
- `month()` 
- `mday()` (day of the month)
- `yday()` (day of the year)
- `wday()` (day of the week)
- `hour()`
- `minute()`
- `second()`

```{r}
datetime <- ymd_hms("2020-03-30 11:59:59")
```

```{r}
year(datetime)
month(datetime)
mday(datetime)
wday(datetime)
wday(datetime, label = TRUE)
```

## Rounding
- `floor_date()` similar to date_trunc() in sql
- `round_date()`
- `ceiling_date()` 

## Time Spans
- *Durations* exact in seconds
- *Periods* weeks or months 
- *Intervals* starting and ending point of time

```{r}
(my_age <- today() - ymd(19940330))
```

**lubridate** provides an alternative that always uses seconds—the duration:
```{r}
as.duration(my_age)
```

A bunch of convenient constructors:
```{r}
dminutes(1)
dyears(1)
```

## Periods
Number of friendly constructor functions

```{r}
seconds(15)
days(30)
years(100)
```

## Time Zones

```{r}
Sys.timezone()
today(tz = "Asia/Ho_Chi_Minh")
```

