---
title: ggplot2
author: Thinh
date: '2020-03-28'
slug: ggplot2
categories: ['bootcamp']
tags:
  - R
backtotop: yes
---
# Data.frame
- columns = variables
- rows = observations 
- cells = values

```{r}
# prepare data to replicate the R bootcamp environment in https://r-bootcamp.netlify.com/chapter1
library(gapminder)
library(tidyverse)
gap1992 <- gapminder %>% filter(year == 1992)
```

Let's start to summary the data 
```{r}
head(gap1992)
```
```{r}
colnames(gap1992)
nrow(gap1992)

```

# Mapping variables to produce geometric plots
A graph consists:

- A `mapping` of variables in `data` to
- `aes()`thetic attributes of
- `geom_`etric objects.
Translated as

```{r}
ggplot(data = gap1992,
       mapping = aes(
         x = log(gdpPercap), 
         y = lifeExp,
         color = continent
         )) + 
  geom_point()
```

**List of aesthetics:**

- x
- y
- size
- shape
- color
- alpha

The great thing about `ggplot2` is that it is easy to swap the presentations.

```{r}
ggplot(data = gap1992,
       mapping = aes(
         x = log(gdpPercap),
         y = lifeExp,
         color = continent
       )) +
  geom_line()
```

The nice thing though is `geoms` are layers on a ggplot. Apply both `geom_point` and `geom_line` at the same time
```{r}
ggplot(data = gap1992,
       mapping = aes(
         x = log(gdpPercap),
         y = lifeExp,
         color = continent
       )) +
  geom_point() + 
  geom_line()
```

Now complete the full plot as show in the bootcmap 
```{r}
ggplot(data = gap1992,
       mapping = aes(
         x = log(gdpPercap),
         y = lifeExp,
         color = continent,
         size = pop
       )) + 
  geom_point() +
  ggtitle("Gapminder for 1992")
```
