---
title: Modelling Building
author: thinh
date: '2020-05-07'
slug: modelling-building
categories:
  - rfds
tags:
  - Book
comments: no
cover: ''
imgs: []
justify: no
lastmod: '2020-05-07T23:05:25+07:00'
license: ''
readingTime: yes
single: no
toc: yes
editor_options: 
  chunk_output_type: console
---
Prerequisite

```{r}
library(tidyverse)
library(modelr)
library(lubridate)
options(na.action = na.warn)
theme_set(theme_light())

```

```{r}
library(nycflights13)
```

# How are low-quality diamonds more expensive? 

```{r}
diamonds %>% 
  mutate(cut = fct_reorder(cut, price)) %>% 
  ggplot(aes(cut, price)) + 
  geom_boxplot() + 
  ggtitle(label = "",
          subtitle = "The worst cut is Fair")
```

```{r}
diamonds %>%
  mutate(color = fct_reorder(color, price)) %>% 
  ggplot(aes(color, price)) + 
  geom_boxplot() + 
  ggtitle(label = "",
          subtitle = "The worst colour is J (slightly yellow)")
```

```{r}
diamonds %>% 
  mutate(clarity = fct_reorder(clarity, price)) %>% 
  ggplot(aes(clarity, price)) + 
  geom_boxplot() + 
  ggtitle(label = "",
          subtitle = "The worst clarity is I1 (inclusions visible to the naked eye)")
```

# Price and carat

```{r}
diamonds2 <- diamonds %>% 
  filter(carat <= 2.5) %>% 
  mutate(lcarat = log2(carat),
         lprice = log2(price)) 
diamonds2 %>% 
  ggplot(aes(lcarat, lprice)) + 
  geom_hex(bins = 50)
```

Look fairly linear 

```{r}
mod_diamonds <- lm(lprice ~ lcarat, data = diamonds2)

grid <- diamonds2 %>%
  data_grid(carat = seq_range(carat, 20)) %>% 
  mutate(lcarat = log2(carat)) %>%    
  add_predictions(mod_diamonds, "lprice") %>% 
  mutate(price = 2 ^ lprice)

diamonds2 %>% 
  ggplot(aes(carat, price)) +
  geom_hex(bins = 50) + 
  geom_line(data = grid, colour = "red", size = 1) + 
  scale_y_continuous(labels = scales::dollar_format()) +
  labs(
    x = "Size of diamonds (Carat)",
    y = "Price"
  ) 

```

The pattern suggests the larger the carat, the cheaper the price is as there are no diamonds beyond 19K.
Let's verify there is any patterns in the resid plots 

```{r}
diamonds2 %>% 
  add_residuals(mod_diamonds, "lresid") %>% 
  ggplot(aes(lcarat, lresid)) + 
  geom_hex(bins = 50)
```


### A more complicated model 

```{r muptiple linear regression}
mod_diamonds2 <- lm(lprice ~ lcarat+ color + cut + clarity, data = diamonds2)
summary(mod_diamonds2)
```

```{r}
grid <- diamonds2 %>% 
  data_grid(cut, .model = mod_diamonds2) %>% 
  add_predictions(mod_diamonds2)

grid %>% 
  ggplot(aes(cut, pred)) +
  geom_point() 
```



```{r}
diamonds2 <- diamonds2 %>% 
  add_residuals(mod_diamonds2, "lresid2") 

diamonds2 %>%
  ggplot(aes(lcarat, lresid2)) +
  geom_hex(bins = 50)

```


### What effects the number of flights? 

```{r}
daily <- flights %>%
  mutate(date = make_date(year, month, day)) %>% 
  count(date) 

daily %>% 
  ggplot(aes(date, n)) + 
  geom_line()
```

Very strong day of week

```{r}
daily <- daily %>% 
  mutate(week = wday(date, label = TRUE, week_start = 1)) 

daily %>% 
  ggplot(aes(week, n, fill = week)) + 
  geom_boxplot() + 
  labs(
    x = "Day of week", 
    y = "Number of flights",
    title = "The number of flights in 2013", 
    subtitle = "Week start is Monday"
  ) 
```

There are few flights indeed on weekends. One way to remove this strong pattern is to use a model. 

First, display predictions to overlaid the original data 
```{r}
mod <- lm(n ~ week, data = daily) 
summary(mod)

# Add prediction 

grid <- daily %>% 
  data_grid(week) %>% 
  add_predictions(mod, "n")

daily %>% 
  ggplot(aes(week, n, fill = week)) + 
  geom_boxplot() + 
  labs(
    x = "Day of week", 
    y = "Number of flights",
    title = "The number of flights in 2013", 
    subtitle = "Week start is Monday"
  ) +
  geom_point(data = grid, colour = "red", size = 4)
```

Second, visualise the resid

```{r}
daily <- daily %>% 
  add_residuals(mod) 

daily %>% 
  filter(resid < -100)

daily %>% 
  ggplot(aes(date, resid)) + 
  geom_ref_line(h = 0, colour = "red") +
  geom_line(aes(colour = week)) + 
  geom_smooth(se = FALSE, span = 0.2)
```

* Our model fails in June
* As well, sunday and saturday seems to be failed to capture in the model 
* Some days are far from expected 

What we are drawn from the model:
* fewer the flights in Jan, Feb
* more flights in summer 
* use the domain of knowledge to brainstorm more 


### Seasonal Saturday Effect

```{r}
daily %>% 
  filter(week == "Sat") %>% 
  ggplot(aes(date, n )) + 
  geom_line() + 
  geom_point() +
  scale_x_date(
    date_breaks = "1 month", 
    date_labels = "%b"
  )
```

Build a function to arrange dates in seasons

```{r}
term <- function(date) {
  cut(date,
      breaks = ymd(20130101, 20130605, 20130825, 20140101),
      labels = c("spring", "summer", "fall")
  )
}

daily <- daily %>% 
  mutate(term = term(date))

# replot 

daily %>% 
  filter(week == "Sat") %>% 
  ggplot(aes(date, n, colour = term)) + 
  geom_line() + 
  geom_point(alpha = 1/3) +
  scale_x_date(
    date_breaks = "1 month", 
    date_labels = "%b"
  ) + 
  labs(
    x = "Month of Year",
    y = "Number of flights",
    colour = "Seasons"
  ) 
```


Let's look at day of week 

```{r}
daily %>% 
  ggplot(aes(week, n, fill = term )) + 
  geom_boxplot()
```

It seems that there is a significant variations across the terms. We might include the terms to improve the model. Let's compare the first 

```{r}
mod1 <- lm(n ~ week, data = daily)
mod2 <- lm(n ~ week*term, data = daily)
```

```{r}
daily %>% 
  gather_residuals(without = mod1, with = mod2) %>% 
  ggplot(aes(date, resid, color = model)) + 
  geom_line()
```

Let's look the performance of model by overlaying the predictions onto the raw data

```{r}
grid <- daily %>% 
  data_grid(week, term) %>% 
  add_predictions(mod2, "n")

daily %>% 
  ggplot(aes(week, n, colour = term)) + 
  geom_boxplot() + 
  geom_point(data = grid, colour = "red") + 
  facet_wrap(~ term)
```

It shows clearly that we fail to capture the data in Fall as the actual data is far from from what the model suggested 


