---
title: Relational Data With dplyr
author: thinh
date: '2020-04-06'
slug: relational-data-with-dplyr
categories:
  - rfds
tags:
  - R
comments: no
cover: ''
imgs: []
justify: no
license: ''
readingTime: yes
single: no
---
# Introduction 
Three families of verbs to work with relational data:

- *Mutating Join*
- *Filtering Join*
- *Set Operation*

Most common relational data is in a RDBMS. SQL is used to manipulate these system. `dplyr` does the same thing but easier 
 
# Prerequisites

```{r}
library(tidyverse)
```


# nycflights 
```{r}
library(nycflights13)
```

Contain 4 tibbles which are related to `flights`: `airlines`, `airports`, `planes`, `weather`
For nycflights13:

• `flights` connects to `planes` via a single variable, tailnum.
• `flights` connects to `airlines` through the carrier variable.
• `flights` connects to `airports` in two ways: via the origin and dest variables.
• `flights` connects to `weather` via origin (the location), and year, month, day, and hour (the time).
 
```{r}
head(airlines,10)
```

```{r}
head(airports,10)
```

```{r}
head(planes,10)
```

```{r}
head(weather)
```

# Keys
- Variables to connect each pair of tables --> called *keys*.
- May require one or more variables to connect each pair of tables. 
- Two type of *keys*:
  - *primary key*: uniquely identifies an observation in **its own table**. `planes$tailnum` in this case
  - *foregin key*: uniquely identifies an observation **in another table**. `flights$tailnum` in this case
- It's good practice to verify if indeed identify unique observation or not 

```{r}
planes %>% 
  count(tailnum) %>% 
  filter(n > 1) # not a unique observation
```

Sometimes, a table doesn't have a explicit primary key such as `flights`

```{r}
flights %>% 
  count(year, month, day, flight) %>% 
  filter(n > 1)
```

If a table lacks a primary key, it’s sometimes useful to add one with `mutate()` and `row_number()`. That makes it easier to match observations if you’ve done some filtering and want to check back in with the original data. This is called a *surrogate key*.


# Mutating Joins 
- Matches observations by their keys, then copies across variables from one table to the other.
- To make it visible, narrow down the table first 
```{r}
flights2 <- flights %>% 
  select(year:day, hour, origin, dest, tailnum, carrier)

flights2
```

- want to add full name of the airline then 

```{r}
flights2 %>% 
  select(-c(origin, dest)) %>% 
  left_join(airlines, by = "carrier")
```

# Understanding joins 

```{r}
x <- tribble( 
  ~key, ~val_x, 
  1, "x1", 
  2, "x2",
  3, "x3" 
  )

y <- tribble(
  ~key, ~val_y,
  1, "y1",
  2, "y2",
  4, "y3"
)
```


## Inner Join
Matches pairs of observations whenever their keys are equal

```{r}
x %>% 
  inner_join(y, by = "key")
```

## Outer Join 
Keeps observations that appear in at least one of the tables. 
There are three types of outer joins:

- *left join* : keep all observation in x 
- *right join*: keep all observation in y
- *full join*: keep all observation in x and y

## Defining the Key Columns
- The default, `by = NULL`, use all variables that are appear in both tables, *natural join*
- `by = "x"` use only some common variables 
- `by = c("a" = "b")` match `variable a` in `table x` with `variable b` in `table y`. For example 

```{r}
flights2 %>% 
  left_join(airports, c("dest" = "faa")) # map destination
```

```{r}
flights2 %>% 
  left_join(airports, c("origin" = "faa"))
```


# Filtering Joins
Match observations in the same way as mutating joins, but affect the observations, not the variables. There are two types

- *semi_join(x, y)* keeps all observation in x that have a match in y
- *anti_join(x,y)* drop all observation in x that a match in y
For examples,
```{r}
top_dest <- flights %>% 
  count(dest, sort = TRUE) %>% 
  head(10)
top_dest
```

We want to find which flight that went to one of these destinations

```{r}
flights %>% 
  filter(dest %in% top_dest$dest)
```

It's difficult when we have multille variables in the filter. To save our time, we can apply `semi_join`

```{r}
flights %>% 
  semi_join(top_dest, by = "dest")

```

Inverse of a `semi_join` is `anti_join`. How many flights that don't have a match with `planes`

```{r}
flights %>% 
  anti_join(planes, by = "tailnum") %>% 
  count(tailnum, sort = TRUE)
```

# Set Operation 
- `intersect(x, y)` return only variables in both x, y
- `union(x, y)` return unique observation in x and y 
- `setdiff(x, y)` return observation in x but not in y 

Simple data
```{r}
df1 <- tibble(
  x = 1:2,
  y = c(1,1)
)

df2 <- tribble(
  ~x, ~y, 
  1, 1,
  1, 2
)
```

```{r}
# intersect
intersect(df1, df2)
```

```{r}
# union 
union(df1, df2)
```

```{r}
# setdiff
setdiff(df1, df2)
```

