---
title: Strings with stringr
author: thinh
date: '2020-04-07'
slug: strings-with-stringr
categories:
  - rfds
tags:
  - R
comments: no
cover: ''
imgs: []
justify: no
lastmod: '2020-04-07T22:20:48+07:00'
license: ''
readingTime: yes
single: no
---

# Prerequisites
```{r}
library(tidyverse)
library(stringr)
```

## String basics 
Create a string can be done with double quotes or single quote

```{r}
string1 <- "thinh"
string2 <- "thinh's code"
```

Escape special characters with backslash

```{r}
(double_quote <- "\"" )# or '"'
```
```{r}
(single_quote <- '\'' ) # or "'"
```

## String Length 
```{r}
str_length(c("a", "thinh tran", NA))
```

## Combining Strings 

**combine**

```{r}
str_c("x", "y" ) # can add sep =
```

**collapse**

```{r}
str_c(c("x","y"), collapse = ",")
```

## Subsetting Strings
Extract part of strings `str_sub()`

```{r}
x <- c("Apple", "Banana", "Orange")
str_sub(x, 1, 3)
```

```{r}
x <- c("Apple", "Banana", "Orange")
str_sub(x, -3, -1)
```


```{r}
str_sub(x, 1, 1) <- str_to_lower(str_sub(x, 1, 1))
x
```


# Matching Pattern With Regular Expression 
## Basic Matches 
Simple

```{r}
x <- c("banana", "apple", "pear")
str_view(x, "an")
```

Match any characters except a new line with `.`

```{r}
str_view(x, ".a.")
```

```{r}
str_view(c("abc", "a.c", "bef"), "a\\.c")
```


## Anchors
*Anchor* the regular expression so that it matches from the start or end of the string. 

- `^` to match the start of string 
- `$` to match the end of string 

```{r}
x <- c("apple", "banana", "pear")
str_view(x, "^a")
```

```{r}
str_view(x, "r$")
```

## Character Classes And Alternatives
- `\d` matches any digits
- `\s` matches any whitespaces
- `[abc]` matches any a,b or c
- `[^abc]` matches any except a,b or c

Alternation for expression

```{r}
str_view(c("thinh", "thịnh" ), "th(i|ị)nh")
```

## Repetition 
How many times pattern match 
- `?`: 0 or 1
- `+`: 1 or more
- `*`: 0 or more

```{r}
x <- "1888 is the longest year in Roman numerals: MDCCCLXXXVIII" 
str_view(x, "CC?") # 0 or 1 CC 
``` 

```{r}
str_view(x, "CC+") # 1 or more (greedy map)
```

```{r}
str_view(x, "C[LX]+") # CLX or CL or CX --> 1 or more X or L 
```

Can also use number of matches 
- `{n}` exactly n 
- `{n, }` n or more
- `{, m}` at most m
- `{n, m}` between n and m

```{r}
str_view(x, "C{2}")
```

```{r}
str_view(x, "C{2,}") # greedy match
```

```{r}
str_view(x, "C{2,}?") # lazy match
```

## Grouping and Backreferences 
```{r}
str_view(fruit, "(..)\\1", match = TRUE) # match any 2-character length repeated once
```

## Tools 
 - Determine which strings match a pattern.
 - Find the positions of matches.
 - Extract the content of matches.
 - Replace matches with new values.
 - Split a string based on a match.
 
### Detect Matches 
use `str_detect()`
```{r}
x <- c("apple", "banana", "pear")
str_detect(x, "an")
```

```{r}
# how many common words start with t
sum(str_detect(words, "^t"))
```

```{r}
# match a pattern
words[str_detect(words, "x$")]
```
Or we can use
```{r}
str_subset(words, "x$")
```

How many matches are in the string? 

```{r}
str_count(x, "a")
```

## Extract Matches
```{r}
stringr::sentences %>% head(10)
```

*Question:* How many sentences containing a colour?

```{r}
colours <- c("red", "orange", "yellow", "green", "blue", "purple")
colour_match <- str_c(colours, collapse = "|")
```

```{r}
has_colour <- str_subset(sentences, colour_match)
(matches <- str_extract(has_colour, colour_match))
```

```{r}
more <- sentences[str_count(sentences, colour_match) > 1]
str_view_all(more, colour_match)
```

