---
title: "Tidy Data with tidyr"
author: "thinh"
date: '2020-04-05'
cover: ''
comments: no
imgs: []
justify: no
lastmod: '2020-04-05T20:18:57+07:00'
license: ''
readingTime: yes
single: no
slug: tidy-data-with-tidyr
categories:
  - rfds
tags:
  - R
output: 
  html_document: 
    theme: cosmo
    toc: yes
---

# What is a tidy data 
**3 rules**:

- each variable has its own column
- each observation has its own row
- each value has its own cell

```{r}
library(tidyverse)
```

```{r}
# this is a tidy data
table1
```

```{r}
# this isn't a tidy data
table2
```

# Deal with untidy data with spread() and gather()
**Problems**:

- 1 variable might spread across multiple columns
- 1 observation might be scatterred across multiple rows

## Gathering
Some column names are not name of variables but *value* of the variable 

```{r}
# 1999 and 2000 are values of year variable
table4a # cases
table4b # population 
```

So we need to gather these columns into a new variable. The process is followed

- set of columns that represent values
- name of variable to form the column name, `key`, in the case `year`
- name of variable to gather, `value`, in the case `case`

```{r}
table4a <- table4a %>% 
  gather('1999', '2000', key = 'year', value = 'cases')
table4a
```

```{r}
table4b <- table4b %>% 
  gather("1999", "2000", key = "year", value = "population")
table4b 
```

Join these tables to achieve `table1`

```{r}
left_join(table4a, table4b)
```


## Spreading
An observation is scatterred across multiple rows. 

```{r}
table2
```

In this table, the observation is a country in year but each spreads two rows. The process to tidy up:

- The column that contains variable names, the `key` column. Here is `type`
- The column that contains values forms multiple columns, the `value` column. Here is `count`

```{r}
table2 %>% 
  spread(key = type, value = count)
```

In summary, `gather()` makes table narrower and longer where as `spread()` makes table wider and shorter.

# Separating and Pull
## separate()
```{r}
# rate make this table untidy
table3
```

Separate rate into `case` and `population`

```{r}
table3 %>% 
  separate(rate, into = c("case", "population"))
```

By default, `separate()` will split values whenever it sees a non-alphanumeric character. We can specify the split then 

```{r}
# add sep = 
table3 %>% 
  separate(rate, into = c("case", "population"), sep = "/")
```

However, this still needs to improves as `case` and `population` are in character type. They shoud be `int`

```{r}
table3 %>% 
  separate(rate , 
           into = c("rate", "population"),
           convert = TRUE)
```

## unite()
- inverse of `separate()`
```{r}
table5
```

**Problem**: century and year can be united to form as year (make more sense)

```{r}
table5 %>% 
  unite(new, century, year)
```
Still require improvement with `underscore _` (by default `sep = _`)

```{r}
table5 %>% 
  unite(new, century, year, sep = "")
```

# Missing Values
A value can be missing in one of two possible ways:

- *explicitly*: flagged with NA
- *implicitly*: not present in the data

```{r}
stocks <- tibble(
  year = c(rep(2015,4), rep(2016,3)),
  qtr = c(seq(1,4), seq(2,4)),
  return = c(1.88, 0.59, 0.35, NA, 0.92, 0.17, 2.66)
)
stocks
```
So, explicit is value of return in 2015, quarter 4. Implicit is the observation of year 2016, quarter 4.

One approach to make the implicit missing to explicit is to put years into columns 

```{r}
stocks %>% 
  spread(year, return)
```

Another way is to use `complete()`

```{r}
stocks %>% 
  complete(year, qtr) # fill missing with NA by distinct all combinations
```

Another method is to fill() with the last observation carried forward
```{r}
treatment <- tribble(
  ~ person, ~ treatment, ~response, 
  "Derrick Whitmore", 1, 7, 
  NA, 2, 10,
  NA, 3, 9, "Katherine Burke", 1, 4 
  )
treatment
```

```{r}
treatment %>% 
  fill(person)
```

# Case study 
Bring everything we learn together 

```{r}
tidyr::who
```

- Firstly, `iso2`, `iso3` are redundant because they convey the same thing with `country`
- Given the structure of columns, `new_sp_`, these are values of the variables, not the names of them.

```{r}
who1 <- who %>% 
  gather(
    new_sp_m014:newrel_f65,
    key = "key",
    value = "cases",
    na.rm = TRUE
  )
who1
```

We can guess some hints of the structure of the data by counting them 

```{r}
who1 %>% 
  count(key)
```

Make variable names consistent 
```{r}
who2 <- who1 %>% 
  mutate(key = stringr::str_replace(key, "newrel", "new_rel"))
who2
```

Then we separate the value in each code (key)

```{r}
who3 <- who2 %>% 
  separate(key, c("new", "type", "sexage"), sep = "_")
who3
```

Zoom in 
```{r}
who3 %>% 
  count(new)
```

So all is new, then drop this redundant variable, together with `iso2`, `iso3`

```{r}
who4 <- who3 %>% 
  select(-c(iso2, iso3, new))
```

Next, separate `sexage` by the first character ( in this case `a`)

```{r}
who5 <- who4 %>% 
  separate(sexage, c("sex", "age"), sep = 1)

who5
```

In the end we only need to do all once 

```{r}
who %>% 
  gather(code, cases, new_sp_m014:newrel_f65, na.rm = TRUE) %>% 
  mutate(
    code = stringr::str_replace(code, "newrel", "new_rel")) %>% 
  separate(code, c("new", "type", "sexage")) %>% 
  select(-c(iso2, iso3, new)) %>% 
  separate(sexage, c("sex", "age"), sep = 1)
```

