---
title: "Data enginnering and data shaping"
author: "Thinh"
date: "6/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(data.table)
```

# Data selection 
## Subsetting rows and columns

```{r}
summary(iris)

iris %>% 
  ggplot(aes(x = Petal.Length, y = Petal.Width, shape = Species, color = Species) ) +
  geom_point(size =2 ) + 
  ggtitle("Petal dimensions by iris species: all measurements")
```

### base R
```{r}
iris[iris$Petal.Length > 2, c("Petal.Length", "Petal.Width", "Species"), drop = FALSE]
```

### data.table 

```{r}
iris_table <- data.table::as.data.table(iris)
iris_table[iris_table$Petal.Length > 2, c("Petal.Length", "Petal.Width", "Species")]
```

### dplyr

```{r}
iris %>% 
  filter(Petal.Length > 2) %>% 
  select(Petal.Length, Petal.Width, Species)
```


## Removing records with imcomplete data

```{r}
glimpse(msleep)

summary(msleep)
```


### base R 

```{r}
msleep[complete.cases(msleep), , drop = FALSE]
```

### data.table 
``` {r}
msleep_table <- as.data.table(msleep)
msleep_table[complete.cases(msleep_table), ]
```

### dplyr

```{r}
msleep %>% 
  filter(., complete.cases(.))
```

## Ordering rows

```{r}
purchase <- tribble(
  ~day, ~hour, ~n_purchase,
  1 , 9 , 5 ,
  2 , 9 , 3 , 
  2 , 11 , 5 ,
  1 , 13 , 1 ,
  2 , 13 , 3 ,
  1 , 14 , 1
)
```

### base R
```{r}
order_index <- with(purchase, order(day, hour))
purchase_ordered <- purchase[order_index, ,drop = FALSE]
purchase_ordered$running_total <- cumsum(purchase_ordered$n_purchase)
purchase_ordered
```


### data.table 

```{r}
DT_purchase <- as.data.table(purchase)

setorderv(DT_purchase, c("day", "hour"))

DT_purchase[ , running_total := cumsum(n_purchase)]

DT_purchase
```

### dplyr 


```{r}
purchase %>% 
  arrange(day, hour) %>% 
  mutate(sum = cumsum(n_purchase))
```


## Advanced use of ordering

The cumulative sum of sales to be per day

### R base 
```{r}
order_index <- with(purchase, order(day, hour))
purchase_ordered <- purchase[order_index, , drop = FALSE]
data_list <- split(purchase_ordered, purchase_ordered$day)
data_list <- lapply(
  data_list, 
  function(di) {
    di$running_total <- cumsum(di$n_purchase)
    di
  }
)
purchase_ordered <- do.call(base::rbind, data_list)
rownames(purchase_ordered) <- NULL
```

### data.table
```{r}
DT_purchase <- as.data.table(purchase)[order(day, hour),
                                       .(hour = hour,
                                         n_purchase = n_purchase,
                                        running_total = cumsum(n_purchase)),
                                       by = "day"]
DT_purchase
```

### dplyr 

```{r}
purchase %>% 
  arrange(day, hour) %>% 
  group_by(day) %>% 
  mutate(running_total = cumsum(n_purchase)) %>% 
  ungroup()
```

# Basic transform

```{r}
library(datasets)
library(lubridate)
glimpse(airquality)

datestr <-  function(year, month, day) {
  paste(year, month, day, sep = "-")
}
```

### base R

```{r}
airquality_with_date <- airquality
airquality_with_date$date <- with(airquality_with_date, ymd(datestr(1973, Month, Day)))
airquality_with_date[ ,c("Ozone", "date")]
```

### data.table

```{r}
as.data.table(airquality)[,
                          date := ymd(datestr(1973, Month, Day))][, c("Ozone", "date")]
```

### dplyr 

```{r}
airquality %>% 
  mutate(date = make_date(1973, Month, Day)) %>% 
  ggplot(aes(date, Ozone)) + 
  geom_point() +
  geom_line()
```

## Fill missing values

```{r}
library(zoo)
```

### base R

```{r}
airquality_corrected <- airquality_with_date
airquality_corrected$OzoneCorrected <- na.locf(airquality_corrected$Ozone, na.rm = FALSE)
```

### data.table 

```{r}
DT_airquality_corrected <- as.data.table(airquality_with_date)[, 
                     OzoneCorrected := na.locf(airquality_with_date$Ozone, na.rm = FALSE)]
```

### dplyr 

```{r}
airquality %>% 
  mutate(date = make_date(1973, Month, Day)) %>% 
  mutate(OzoneCorrected = zoo::na.locf(Ozone, na.rm = FALSE)) %>% 
  ggplot(aes(date, OzoneCorrected)) +
  geom_point() + 
  geom_line()
```

# Aggregate transforms 
## Combining many rows into summary rows

### Base R

```{r}
iris_summary <- aggregate(
  cbind(Petal.Length, Petal.Width) ~ Species,
  data = iris,
  FUN = mean)

```

### data.table 

```{r}
DT_iris <- as.data.table(iris)

DT_iris[,
        .(Petal.Length = mean(Petal.Length),
          Petal.Width = mean(Petal.Width)), 
        by = .(Species)]
```

### dplyr 

```{r}
iris %>% 
  group_by(Species) %>% 
  summarise(across(c(Petal.Length, Petal.Width), mean)) %>% 
  ungroup()
```


## Window Function 

### Base R

```{r}
iris_copy <- iris

iris_copy$mean_petal_length <- 
  ave(iris_copy$Petal.Length, iris_copy$Petal.Width, FUN = mean)
```

### data.table 

```{r}
DT_iris <- as.data.table(iris)

DT_iris[,
        `:=` (mean_petal_length = mean(Petal.Length),
              mean_petal_width = mean(Petal.Width))]
```


### dplyr

```{r}
iris %>% 
  group_by(Species) %>% 
  mutate(across(c(Petal.Length, Petal.Width), mean, .names = "{col}_mean"))
```


# Multitable data transforms
## Combining two or more ordered data frames quickly 

```{r}
product_table <- tribble(
  ~product_id, ~price, 
  "p1", 9.99, 
  "p2", 16.29,
  "p3", 19.99,
  "p4", 5.49,
  "p5", 24.49
)

sales_table <- tribble(
  ~product_id, ~sold_store, ~sold_online, 
  "p1", 6, 64,
  "p2", 31, 1,
  "p3", 30, 23,
  "p4", 31, 67,
  "p5", 43, 51
)

product_table2 <- tribble(
  ~product_id, ~price,
  "n1", 25.49,
  "n2", 33.99,
  "n3", 17.99
)
```

## Appending Rows 

### base R

```{r}
rbind(product_table, product_table2)
```

### data.table 

```{r}
library(data.table)
rbindlist(list(product_table, product_table2))
```

### dplyr 

```{r}
product_table %>%  
  bind_rows(product_table2)
```

## Splitting tables

### base R 

```{r}
product_table_mask <- product_table
product_table_mask$table <- "product_table"
product_table2_mask <- product_table2
product_table2_mask$table <- "product_table2"

rbind_base <- rbind(product_table_mask, product_table2_mask)

## split 

split(rbind_base, rbind_base$table)
```

### data.table 

```{r}
dt <- as.data.table(rbind_base)

dt[ , max_price := max(price), by = eval("table")]

dt
```


### dplyr 

```{r}
rbind_base %>% 
  group_by(table) %>% 
  mutate(max_price = max(price)) %>% 
  ungroup()
```


## Appending columns 

### base R
```{r}
cbind(product_table, sales_table[, -1])
```

### data.table 

```{r}
cbind(as.data.table(product_table), 
      as.data.table(sales_table[, -1]))
```

### dplyr 

```{r}
product_table %>% 
  bind_cols(sales_table[, -1])
```


## Principal methods to combine from multiple tables
 
### left join 

### base R

```{r}
merge(product_table, sales_table, by = "product_id", all.x = TRUE)
```
 
### data.table 

```{r}
dt_product_table <- as.data.table(product_table)
dt_sales_table <- as.data.table(sales_table)

dt_sales_table[dt_product_table, on = "product_id"]

merge(dt_product_table, dt_sales_table)
```

### dplyr

```{r}
product_table %>% 
  left_join(sales_table)
```


# Reshaping Transform 

Moving data between rows and columns - pivoting 

## Moving data from wide to tall 

*Data example*

```{r}
library(datasets)
library(lubridate)

# clean
seatbelts <- as_tibble(Seatbelts) %>% 
  mutate(dates = index(as.xts(time(Seatbelts))),
         date = make_date(year(dates), month(dates), 01))
         
# restrict down to 1982 to 1983

seatbelts_tidy <- seatbelts %>% 
  filter(between(year(date), 1982, 1983)) %>% 
  mutate(law = ifelse(law == 1, "new law", "pre-law")) %>% 
  select(date, DriversKilled, front, rear, law)

seatbelts_tidy %>% 
  ggplot(aes(date, DriversKilled, colour = law)) + 
  geom_point() +
  geom_smooth(se = FALSE) + 
  ggtitle("UK car drivers death by month")

```

**Questions**: how different between the seat positions in term of death? 

### solution 1: `data.table::melt.data.table()`

```{r}
seatbelts_tidy %>% 
  as.data.table() %>% 
  melt.data.table(id.vars = NULL,
                  measure.vars = c("DriversKilled", "front", "rear"),
                  variable.name = "victim_type",
                  value_name = "nvictims")
```


### solution 2: `tidyr`

```{r}
seatbelts_tidy %>% 
  pivot_longer(c(DriversKilled, front, rear), names_to = "features", values_to = "nvictims") %>% 
  ggplot(aes(date, nvictims, colour = law)) + 
  geom_point() + 
  geom_smooth(se = FALSE) + 
  facet_wrap(~features, scales = "free_y", ncol = 1)
```

## Moving data from tall to wide 

```{r}
chickweight <- ChickWeight %>% 
  as_tibble() %>% 
  select(-Diet) %>% 
  mutate(Chick = str_c("Chick", as.character(Chick)))
  
```


### **solution 1**: `data.table::dcast.data.table()`

```{r}
chickweight %>%
  as.data.table() %>% 
  dcast.data.table(
    Chick ~ Time, 
    value.var = "weight"
  )
```


### **solutionn 2** `tidyr`

```{r}
chickweight %>% 
  pivot_wider(names_from = Time, values_from = weight)
```

