---
title: "Exploring Data"
author: "Thinh"
date: "6/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
library(tidyverse)
theme_set(theme_light())
```

```{r}
cusdata <- read_rds("data/custdata.RDS") %>% as_tibble()

cusdata
```

# Spotting the problems with a summary 

```{r}
skimr::skim(cusdata)
```

Summary statistics can help to spot the problems in the first place 

* Min(`income`) is negative so there is a potential problem. 
* `is_employed` is a-third missing values
* `age` range is from 0 to 120 which is questionable.
* `income` has a wide range. Also the data should not be to narrow.

# Spotting the problems using graphics and visualisation 
* histogram
* density 
* bar charts
* scatter plots and smoothing curves

```{r}
cusdata %>% 
  ggplot(aes(gas_usage)) + 
  geom_histogram(binwidth = 10) +
  labs(
    x = "Gas usage",
    y = "Frequency",
    title = "The distribution of gas usage")
```


```{r}
cusdata %>% 
  ggplot(aes(income)) + 
  geom_density() +
  scale_x_log10(labels = scales::dollar, breaks = c(10, 100, 1000, 10000, 100000, 1000000)) +
  scale_y_continuous(labels = scales::percent_format()) + 
  annotation_logticks(sides = "bt", colour = "gray") + 
  labs(
    x = "Income",
    y = "Density",
    title = "The distribution of income"
  )
```


```{r}
cusdata %>% 
  count(marital_status, sort = TRUE) %>% 
  ggplot(aes(marital_status, n)) + 
  geom_col() + 
  labs(
    x = "Marital status", 
    y = "# Customers"
  )
```


```{r}
cusdata %>% 
  count(state_of_res, sort = TRUE) %>% 
  mutate(state_of_res = fct_reorder(state_of_res, n)) %>%  
  ggplot(aes(state_of_res, n, fill = state_of_res)) + 
  geom_col() + 
  theme(legend.position = "None") + 
  coord_flip() + 
  labs(
    x = "",
    y = "# Customers",
    title = "What is the state of residence of the customers?"
  )
  
```


```{r}
set.seed(2345566)
# reduce the size of data with sampling
sample_custdata<- cusdata %>% 
  sample_frac(size = .1, replace = FALSE) %>% 
  filter(0 < age & age < 100 &
         0 < income & income < 200000)
  
sample_custdata %>% 
  ggplot(aes(age, income)) + 
  geom_point(alpha = .1) +
  geom_smooth() +
  labs(
    x = "Age",
    y = "Income",
    title = "The relationship between ages and income"
  )
```


```{r}
sample_custdata %>% 
  ggplot(aes(age, income)) +
  geom_hex(alpha = .8) + 
  geom_smooth(colour = "black")
```


```{r}
cusdata %>% 
  count(marital_status, health_ins, sort = TRUE) %>% 
  ggplot(aes(marital_status, n, fill = health_ins)) + 
  geom_col() + 
  labs(
    x = "Martial status",
    y = "# Customers",
    fill = "Health insurance",
    title = "Health insurance versus marital status"
  )
```

```{r}
cusdata %>% 
  count(marital_status, health_ins, sort = TRUE) %>% 
  ggplot(aes(marital_status, n, fill = health_ins)) + 
  geom_col(position = position_dodge()) + 
  labs(
    x = "Martial status",
    y = "# Customers",
    fill = "Health insurance",
    title = "Health insurance versus marital status"
  )
```


```{r}
cusdata %>% 
  count(marital_status, health_ins, sort = TRUE) %>% 
  ggplot(aes(marital_status, n, fill = health_ins)) + 
  geom_col(position = position_stack()) + 
  labs(
    x = "Martial status",
    y = "# Customers",
    fill = "Health insurance",
    title = "Health insurance versus marital status"
  ) + 
  scale_fill_gra
```


```{r}
cusdata %>% 
  filter(!is.na(housing_type)) %>% 
  count(marital_status, housing_type, sort = TRUE) %>% 
  mutate(housing_type = fct_reorder(housing_type, n),
         marital_status = fct_reorder(marital_status, n)) %>% 
  ggplot(aes(housing_type, n, fill = marital_status)) +
  geom_col(position = position_dodge()) + 
  scale_fill_brewer(palette = "Dark2") +
  coord_flip() +
  theme(legend.position = "None") +
  facet_wrap(~ marital_status, ncol = 2, scales = "free_x") +
  labs(
    x = "# Customers",
    y = "Housing Type",
    title = "The distribution of marital status by housing type"
  ) 
```

```{r}
sample_custdata %>% 
  ggplot(aes(age)) +
  geom_histogram() + 
  facet_wrap(~marital_status, scales = "free_x")
```

