---
title: "Model Diagnostics"
author: "Thinh"
date: "7/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
```

```{r}
housing <- read_csv("http://www.jaredlander.com/data/housing.csv")

names(housing) <- c("Neighborhood", "Class", "Units", "YearBuilt",  "SqFt", "Income", "IncomePerSqFt", "Expense", "ExpensePerSqFt", "NetIncome", "Value", "ValuePerSqFt", "Boro")
```

Brooklyn and Queens make up one mode and Manhattan makes up the other, while there is not much data on the Bronx and Staten Island

```{r}
housing %>% 
  ggplot(aes(ValuePerSqFt, fill = Boro)) +
  geom_histogram(binwidth = 10) + 
  labs(x = "Value per square foot") + 
  facet_wrap(~ Boro)
```


```{r}
housing %>% 
  filter(Units < 1000) %>% 
  ggplot(aes(SqFt)) + 
  geom_histogram()

housing %>% 
  filter(Units < 1000) %>%
  ggplot(aes(Units)) + 
  geom_histogram()

gridExtra::grid.arrange()
```
There are few outliers beyond 1000. It's reasonable to drop them out? Let see some more information 

```{r}
housing %>% 
  ggplot(aes(SqFt, ValuePerSqFt, colour = Units)) + 
  geom_point() + 
  scale_x_continuous(labels = scales::comma) +
  scale_colour_gradient2(low = "green", high = "red", mid = "white", midpoint = 1000)
```


```{r}
housing %>% 
  ggplot(aes(Units, ValuePerSqFt, colour = Units)) + 
  geom_point() + 
  scale_x_continuous(labels = scales::comma) +
  scale_colour_gradient2(low = "green", high = "red", mid = "white", midpoint = 1000)
```

It's reasonable to drop them out. 


# Modelling 

```{r}
house1 <- housing %>% 
  lm(ValuePerSqFt ~ Units + SqFt + Boro, data = .)

summary(house1)
broom::tidy(house1)
```

```{r}
coefplot::coefplot(house1)
```



# How do we judge the quality of a model? 
This could be an analysis of residuals, the results of an ANOVA test or a Wald test, drop-in deviance, the AIC or BIC score, cross-validation error or bootstrapping

## Residuals 

The difference between actual response and the fitted values. 
The idea is if the model is appropriately fitted to the data, the residuals should be **normally distributed**

## Residuals versus Fitted 
```{r}
h1 <- augment(house1) %>% 
  ggplot(aes(.fitted, .resid)) + 
  geom_point() + 
  geom_hline(yintercept = 0, colour = "grey") + 
  geom_smooth(se = FALSE) + 
  labs(
    x = "Fitted Value",
    y = "Residuals",
    title = "Plot of residuals versus fitted values for house1"
  )

h1
```

It's clearly not random.

```{r}
h1 +
  geom_point(aes(color = Boro))
```


## QQ Plot

```{r}
augment(house1) %>% 
  ggplot(aes(sample = .std.resid)) + 
  geom_qq() +
  geom_abline()
```

If the model is a good fit, the standardized residuals should all fall along a straight line when plotted against the theoretical quantiles of the normal distribution.

## Histogram of residuals 

```{r}
augment(house1) %>% 
  ggplot(aes(.resid)) + 
  geom_histogram()
```
This does not look normally distributed, meaning our model is incomplete
